---
title: "wcGeneSummary"
author: "Noriaki Sato"
package: wcGeneSummary
output: 
    BiocStyle::html_document:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{wcGeneSummary}
    %\VignetteEngine{knitr::knitr}
    %\VignetteEncoding{UTF-8}  
---

# Basic usage

```{r load, warning=FALSE, message=FALSE}
library(wcGeneSummary)
library(org.Hs.eg.db)
library(ggplot2)
```

## Producing word clouds

The basic usage of the package is producing a word cloud of gene summaries by querying Entrez gene IDs.
As many of the words are commonly observed, you should limit word frequency by `excludeFreq`, which is default to 5000.

```{r basic, fig.cap="A wordcloud of gene summaries.", warning=FALSE, message=FALSE, cache=TRUE}
erccs <- c("ERCC1","ERCC2","ERCC3","ERCC4","ERCC5","ERCC6","ERCC8")
entrezID = AnnotationDbi::select(org.Hs.eg.db, keys = erccs, columns = c("ENTREZID"), keytype = "SYMBOL")$ENTREZID
gwc <- wcGeneSummary(entrezID)
gwc$wc
```

It accepts values of the `wordcloud()` function. `numWords` specifies how many words are to be shown on word clouds.

```{r basic2, fig.cap="A wordcloud of gene summaries, customized visualization", warning=FALSE, message=FALSE, cache=T}
## Use palette from palettetown, palettetown::pokepal(150) (Mewtwo!)
gwc <- wcGeneSummary(entrezID, random.order=FALSE,
                     numWords=100,
                     colors=palette(), rot.per=0.4)
gwc$wc
```

And it also returns a data frame consisting of frequency of each term.

```{r table}
knitr::kable(
  head(gwc$df), caption = 'Term frequencies.',
  row.names = FALSE
)
```

## Producing correlation networks

```{r basic3, fig.cap="A correlation network of gene summaries", warning=FALSE, message=FALSE, cache=T}
library(ggraph)
net <- wcGeneSummary(entrezID, plotType="network")
net$net
```

## Text over represenatation analysis (experimental)

```{r ora}
keggPathways <- org.Hs.egPATH2EG
mappedKeys <- mappedkeys(keggPathways)
keggList <- as.list(keggPathways[mappedKeys])

geneList <- keggList$`00785` # Lipoic acid metabolism
pvs <- textORA(geneList)
hist(pvs)
pvs[order(pvs)] |> head()

geneList <- keggList$`05150` # Staphylococcus aureus infection
pvs <- textORA(geneList)
hist(pvs)
pvs[order(pvs)] |> head()
```

Filter words using ORA threshold and frequency threshold by setting `ora=TRUE`.

```{r oranet, fig.cap="A correlation network of gene summaries (ORA)", warning=FALSE, message=FALSE, cache=T}
net <- wcGeneSummary(entrezID, plotType="network", ora=TRUE, edgeLink=FALSE)
net$net
```

# Custom usage

## Interactive inspection of Bayesian network annotated by words

In this example, a Bayesian network showing the module eigengenes relationship are inferred using `boot.strength` function in `bnlearn` from the weighted gene correlation network analysis (WGCNA) results. The modules are annotated by word clouds produced by `wcGeneSummary()`, and can be exported to the format of `Cytoscape.js` or `vis.js`. In this way, module relationship can be interactively inspected with the functional implications.

```{r mod, warning=FALSE, message=FALSE}
## In this example, we simulate WGCNA results.
## you can just use results from WGCNA.
## Assuming WGCNA results are stored in `mod`
mod <- wcGeneSummary::returnExample()
MEs <- mod$MEs
modColors <- mod$colors
ensg <- names(modColors)

# library(bnlearn)
library(igraph)

## Replace like boot.strength(mod$MEs, R=500, algorithm = "hc")
# dag <- model2network("[ME1][ME2|ME1]") # If using bnlearn
dag <- graph_from_literal( ME1-+ME2, ME1-+ME3 )

## Convert to igraph
# g <- as.igraph(dag)
g <- dag

## Assign edge attributes
## Skip, if you perform boot.strength, the edge attributes can be added from the result
# el <- data.frame(as_edgelist(g))
# colnames(el) <- c("from","to")
# el <- left_join(el, bs)
# E(g)$strength <- el$strength
# E(g)$direction <- el$direction

## Node attributes
V(g)$stripName <- gsub("ME","",V(g)$name)
sizes <- table(modColors)
V(g)$size <- as.numeric(sizes[V(g)$stripName])

## Directory to save images and a script
rootDir <- "./"
netDir <- "visCyjs"
imageDir <- "images"

dir.create(paste0(rootDir, netDir))
dir.create(paste0(rootDir, netDir, "/", imageDir))

images <- c()
plotType <- "bar"
numLim <- 200 # limit for gene number
for (i in V(g)$name){
    print(i)
    i <- as.numeric(gsub("ME","",i)) # strip ME

    queries <- ensg[modColors==i]
    if (length(queries)>numLim) {
        warning("Sampling random genes")
        queries <- queries[sample(1:length(queries), numLim)] ## Temporary restrict to randomly chosen genes, should be replaced to like kME values
    }
    
    ## Convert to ENTREZ
    entre <- AnnotationDbi::select(org.Hs.eg.db, keytype="ENSEMBL", keys = queries, columns = "ENTREZID")$ENTREZID
    
    if (plotType=="bar"){
        plt <- makeBar(entre) # get barplot
    } else { ## If wordcloud
        # A <- wcGeneSummary(entre,
        #                    madeUpper=c("dna","rna",
        #                                tolower(keys(org.Hs.eg.db, keytype="SYMBOL"))))
        # palNum <- sample(1:151,1) # palettetown
        # 
        # ## This time use ggwordcloud()
        # plt <- ggwordcloud::ggwordcloud(A$df$word, A$df$freq,
        #                      shape="circle", min.freq = 10,
        #                      rot.per = 0.5, random.order = FALSE,
        #                      colors = pokepal(palNum))+
        #          scale_size_area(max_size = 40)
    }
    ## Save images
    ggsave(paste0(rootDir, netDir, "/", imageDir, "/", i ,".png"),
           plt, dpi=300, width=10, height=10)
    ## Store image dir
    images <- c(images, paste0(imageDir, "/", i ,".png"))
}
V(g)$image <- images

## Node shape
if (plotType=="bar"){
    V(g)$shape <- rep("rectangle", length(V(g))) 
} else {
    V(g)$shape <- rep("circle", length(V(g)))
}

## Scale the node size
sizeMin <- 50
sizeMax <- 200
rawMin <- min(V(g)$size)
rawMax <- max(V(g)$size)
scf <- (sizeMax-sizeMin)/(rawMax-rawMin)
V(g)$size <- scf * V(g)$size + sizeMin - scf * rawMin

## Export
exportCyjs(g, rootDir, netDir)
# or, exportVisjs(g, rootDir, netDir)
```

Use like `http-server` in the directory containing a exported JavaScript, and interactively inspect the module relationship.

### Annotating module eigengenes dendrogram

```{r wgcna, fig.cap="A ME dendrogram with words", warning=FALSE, message=FALSE, cache=T}
plotEigengeneNetworksWithWords(MEs, modColors)
```

```{r}
sessionInfo()
```