---
title: "Mining microbiome-related literature"
author: "Noriaki Sato"
package: wcGeneSummary
output: 
    BiocStyle::html_document:
    toc_float: true
vignette: >
    %\VignetteIndexEntry{wcGeneSummary}
    %\VignetteEngine{knitr::knitr}
    %\VignetteEncoding{UTF-8}  
---
    
The general usage for mining the information related to microbiome-related signatures.

```{r load, warning=FALSE, message=FALSE}
library(wcGeneSummary)
library(ggplot2)
library(ggraph)
library(RColorBrewer)
```

We use BugSigDB, and its R port bugsigdbr to obtain the curated dataset of the relationship with bacterial taxonomy and human diseases ((Geistlinger et al. 2022)[https://bugsigdb.org/Main_Page]). Users can query microbiome names, which will be searched for MetaPhlAn taxonomic annotation. If `target="title"`, the title of the corresponding articles will be summarized.

```{r bsdb_basic, warning=FALSE, message=FALSE}
basic <- wcBSDB(c("Veillonella dispar","Neisseria flava"),tag=TRUE,
    curate=TRUE,target="title",pre=TRUE,cl=snow::makeCluster(12),
    pal=RColorBrewer::brewer.pal(10, "Set2"),numWords=80)
basic@freqDf
basic@wc
```
If `target="abstract"`, the corresponding abstract will be fetched and be summarized.

```{r bsdb_basic2, warning=FALSE, message=FALSE}
basic2 <- wcBSDB(c("Veillonella dispar","Neisseria flava"),tag=TRUE,
    curate=TRUE,target="abstract",pre=TRUE,cl=snow::makeCluster(12),
    pal=RColorBrewer::brewer.pal(10, "Set2"),numWords=80)
basic2@freqDf
basic2@wc
```
For successful visualization, pre-caculated TF-IDF and frequency data frame is available and one can use them to filter the highly occurring words, or the other prefiltering option used in `wcGeneSummary`.

```{r filters, warning=FALSE, message=FALSE}
rmwords <- wcGeneSummary:::allTfIdfBSDB
filter <- rmwords[rmwords$tfidf>quantile(rmwords$tfidf, 0.95),]
filter
```
The network visualization is possible by enabling `plotType="network"`.

```{r bsdb_basic_network, warning=FALSE, message=FALSE}
net <- wcBSDB(c("Veillonella dispar","Neisseria flava"),
    curate=TRUE,target="title",pre=TRUE,plotType="network",
    additionalRemove=filter$word, corThresh=0.2,
    numWords=40)
net@net
```

The words-to-species relationship can be plotted by `mbPlot=TRUE`.

```{r bsdb_basic_network_mb, warning=FALSE, message=FALSE}
net2 <- wcBSDB(c("Veillonella dispar","Neisseria flava",
    "Veillonella parvula","Akkermansia muciniphila"), mbPlot=TRUE,
    curate=TRUE,target="title",pre=TRUE,plotType="network",
    additionalRemove=filter$word,
    numWords=40, corThresh=0.2, colorText=TRUE)
net2@net
```

As the BugSigDB contains the relationship between bacterial taxonomy and disease, disease name can also be plotted. When `disPlot=TRUE`, the `mbPlot`
 will be set to `TRUE` by default.
 
```{r bsdb_basic_network_dis, warning=FALSE, message=FALSE}
net3 <- wcBSDB(c("Veillonella dispar","Neisseria flava",
    "Veillonella parvula","Akkermansia muciniphila"), mbPlot=TRUE,
    curate=TRUE,target="title",pre=TRUE,plotType="network",
    additionalRemove=filter$word, disPlot=TRUE,
    numWords=40, corThresh=0.2, colorText=TRUE)
net3@net
```

```{r}
sessionInfo()
```